Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2016-11-11 paper/numerical_experiments_v3.tex
stashing edits

The unknown parameters to be inferred correspond to the forcing field, so that in this case $f(q)=q$. The observations are sparse measurements of the state.
\cref{fig:baseSetup} shows the locations of the observations and the region $\Omega_I$ over which the QoI is calculated. For the low-fidelity model, the inverse problem is linear (i.e., the observations depend linearly on the parameters). Since the inverse problem is ill-posed, we use Tikhonov regularization~\cite{EngHanNeu00}; the regularization term is $\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2\:\textrm{d}A$, where $\beta=10^{-5}$ is the regularization coefficient. % KW: term in what equation? give a reference or re-state the objective here.
%
\begin{figure}[htbp]
\centering
\includegraphics[width=0.8\textwidth]{baseSeries/setup_3_3.pdf}
\caption{Locations of the observations and the QoI region.}
\label{fig:baseSetup}
\end{figure}
%

@@

We see that the largest local error contribution is concentrated in the QoI region, and the data point closest to the QoI. 
% KW: I am not sure how to read the second phrase in this first sentence. Is it '...contribution is concentrated in two places: in the QoI region and near the observation location closest to the QoI.'  ??
In the first decomposition of the error (\cref{fig:baseRef0}), the region where the elemental error is maximum is the leftmost data point. % KW: again here, what is a 'data point' ? Do you mean 'observation location' ?
Since the constraining model is an elliptic PDE, with weak convection, information flow is localized, and is weakly convected from left to right. Therefore, for the calculation of the QoI, it is most important to refine the region near the leftmost data point, and the QoI region. % KW: same here. plus another confusing comma.
After that, the error decomposition suggests refinement in regions upstream and around the middle data point, and then the rightmost data point. % KW: and here

@@

%------------------------------------------------------------%
\subsubsection{Interaction of Observations and QoI} \label{sec:qoivdata}
%------------------------------------------------------------%
%
The error estimate decomposition suggests the use of the high-fidelity model in areas of the domain that are important to the interaction between the observations and QoI; the interaction between these two can be complex, and the areas suggested for refinement may be nonintuitive. To see this, we compare the error estimate decomposition for three sizes of the QoI region $\Omega_I$ given the same set of data points, and for three nested sets of data points given the same QoI region. For the sake of illustration, we make two refinement iterations for each combination of observations and QoI region, regardless of the magnitude of the relative error estimate. However, in conducting the numerical experiments, it was observed that the number of iterations needed to achieve a given tolerance tended to increase as the QoI region increased.

\cref{fig:qoiStudy} shows the domain for three cases considered, with the same set of observations but increasingly large, nested QoI regions $\Omega_I$.
The error decompositions for each case are also shown in \cref{fig:qoiStudy}. The bottom row gives the baseline case presented in \cref{sec:cdvcdrBaseRef}, although here we choose the basis functions $i$ whose error $\varepsilon_i$ are among the largest $5\%$, so the proportion of additional refined elements in each iteration is slightly larger.
% KW: I don't understand this last sentence. What's the difference with Section 4.1.2?
Although refinement is still most important around the data point closest to $x_1=0$, as the QoI region expands the other two data points become more important in that the error decomposition suggests refinement around them earlier. As the QoI region expands, it is also more clearly noticeable that refinement is not equally important in all parts of the QoI region.
% KW: fix references to 'data point' throughout.

@@

\cref{fig:dataStudy} shows another set of cases considered, now with the same QoI region $\Omega_I$ but with increasing, nested sets of observations.
The error decomposition for the three cases is shown in \cref{fig:dataStudy}. Again, the bottom row gives the baseline case presented in \cref{sec:cdvcdrBaseRef}, although here the basis functions with the largest $5\%$ of the error are chosen, so the proportion of additional refined elements in each iteration is slightly larger. Refinement appears to be consistently most important around the data point closest to $x_1=0$ and the QoI region. However, as more data points are added, it becomes no longer necessarily true that refinement becomes less important around data points as their distance from the QoI region increases. The data points also interact with each other in that placing data points in regions of previously relatively uniform error contribution tends to result in a new error decomposition that is positive or negative around the data points, with valleys of zero magnitude in between. % KW: I don't understand a lot of this discussion. You need to explain to me what you mean.

@@

We consider the same high-fidelity model as in \cref{sec:cdvcdr}:
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in U,
\end{equation}
with the same diffusion coefficient $k_d = 0.1$  and reaction coefficient $k_r = -42$. The low-fidelity model
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in\R
\end{equation}
differs from the high-fidelity model only in that the parameter $q$ is a constant instead of a field. The intermediate mixed-fidelity models thus have parameter fields that are constant over the subregions of the domain where the low-fidelity model is used. For ease of implementation, we require that the resulting parameter field remain continuous at the interface between the low-fidelity and high-fidelity subdomains, although this constraint is not necessary for the theory to hold. The domain, mesh, boundary conditions, and velocity field, as well as the observations, unknown parameters to be inferred, and QoI, remain the same as described in \cref{sec:cdvcdr}. As the inverse problem is ill-posed, except for perhaps % KW: is it or not?
in the case where the low-fidelity model is used throughout the domain, regularization is added; % KW: added to what?
 the Tikhonov regularization term is $\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2+f(q)^2\:\textrm{d}A$, where $\beta=10^{-3}$ is a regularization coefficient. % KW: refer to the actual equation you are adding regularization to, or re-state the objective.

@@

\Cref{fig:svfRef} shows the local error contributions, as well as the subdomains where the low- and high-fidelity models are used, for the first two and last mixed-fidelity model thus generated. Each linear Lagrange basis function's contribution is plotted at its nonzero node. % KW: what does this sentence mean and is it really necessary?
%
\begin{figure}[htbp]
\centering
\subfloat[MF$_0$ ($0\%$ HF)]{
	\includegraphics[width=0.46\textwidth]{svf/cd_cdr_LF_divvy.png}
  \includegraphics[width=0.49\textwidth]{svf/err_breakdown_LF.png}
} \\
\subfloat[MF$_1$ ($10\%$ HF)]{
	\includegraphics[width=0.46\textwidth]{svf/cd_cdr_MF01_divvy.png}
  \includegraphics[width=0.49\textwidth]{svf/err_breakdown_MF01.png}
} \\
\subfloat[MF$_2$ ($20\%$ HF)]{
  \includegraphics[width=0.46\textwidth]{svf/cd_cdr_MF02_divvy.png}
  \includegraphics[width=0.49\textwidth]{svf/err_breakdown_MF02.png}
} \\
%\begin{subfigure}[b]{\textwidth}
%	\centering
%	\includegraphics[width=0.48\textwidth]{svf/cd_cdr_MF03_divvy.png}
%  \includegraphics[width=0.51\textwidth]{svf/err_breakdown_MF03.png}
%  \vspace{-0.7\baselineskip}
%  \caption{MF$_3$ ($30\%$ HF)}
%  \vspace{0.8\baselineskip}
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}
%	\centering
%	\includegraphics[width=0.48\textwidth]{svf/cd_cdr_MF04_divvy.png}
%  \includegraphics[width=0.51\textwidth]{svf/err_breakdown_MF04.png}
%  \vspace{-0.7\baselineskip}
%  \caption{MF$_4$ ($40\%$ HF)}
%  \vspace{0.8\baselineskip}
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}
%	\centering
%	\includegraphics[width=0.48\textwidth]{svf/cd_cdr_MF05_divvy.png}
%  \includegraphics[width=0.51\textwidth]{svf/err_breakdown_MF05.png}
%  \vspace{-0.7\baselineskip}
%  \caption{MF$_5$ ($50\%$ HF)}
%  \vspace{0.8\baselineskip}
%\end{subfigure}
\subfloat[MF$_6$ ($60\%$ HF)]{
	\includegraphics[width=0.46\textwidth]{svf/cd_cdr_MF06_divvy.png}
  \includegraphics[width=0.49\textwidth]{svf/err_breakdown_MF06.png}
}
\caption{Left: Multi-fidelity refinement over the domain (low-fidelity constant-parameter model used in red portion, high-fidelity field-parameter model used in blue portion). Right: local error contributions. The (weighted) residual, and thus the local error contribution, tends to spike sharply at the interface between the low- and high-fidelity regions; the color range is truncated to make the error distribution visible elsewhere in the domain.}
\label{fig:svfRef}
\end{figure}
%
Comparing to \cref{fig:baseRef}, we see that in this case the local error contribution is not as greatly concentrated around the QoI region and the nearest data point; here, all three data points and the QoI region have associated regions of sufficiently similar high local error that all are refined in the first iteration. This reflects the global nature of the differences between the low- and high-fidelity models. % KW: fix data-point language

@@

The unknown parameters we wish to infer correspond to the source term $f(q)=q$; we impose $f(q)=q=0$ on the boundary $\partial\Omega$. Synthetic observations at 18 points in the domain are generated by running the high-fidelity model on a finer mesh. The locations of the observations as well as the QoI region $\Omega_I$ are shown in \cref{fig:setup3D}. We use a regularization term $\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2\:\textrm{d}V$. % KW: again please connect reg term to an equation or just state objective function here.
%
\begin{figure}[htbp]
\centering
\includegraphics[width=0.4\textwidth]{series3D/setup_aerial_nolegend.pdf} \hfill
\includegraphics[width=0.55\textwidth]{series3D/setup_3view.pdf} \\
\vspace{\baselineskip}
\includegraphics[width=0.6\textwidth]{series3D/setup_side_view.pdf}
\caption{Three views of the locations of the observations and the QoI region.}
\label{fig:setup3D}
\end{figure}
%

@@

We now present the results of solving the inference problem using \cref{alg:refSeries}, with a relative error tolerance of $0.1\%$. At each iteration, we choose the $2\%$ of the basis functions with the largest error for model refinement; since each linear Lagrange basis function has eight elements in its support, the number of additional elements marked for refinement in each iteration is usually greater than for the 2D examples. All simulations are run on a single processor; we use the default nonlinear solver in \texttt{libMesh} (Newton's method with Brent line-search), and linear solves are performed using PETSC's GMRES solver, preconditioned by incomplete factorization. % KW: need to add a citation to papers for libMesh and PETSC?

@@

% KW: this paragraph is a bit rambling. I think it is obvious: you do better when you optimize the model for the actual dataset. I think we leave this out unless you think there is a critical point here that is not captured elsewhere.
%In testing this online-offline setting with a larger range of datasets, we would note some additional observations. Suppose one has two datasets $d_1$ and $d_2$, and wishes to perform inference on $d_2$. For a fixed proportion of domain refinement, a mixed-fidelity model that was adaptively generated using $d_2$ gives a smaller QoI error than a mixed-fidelity model that was generated based on $d_1$. The performance gap tends to widen as the two datasets become more different. Suppose one generated $d_1$ using the high-fidelity model $d_2$ using a model which differed from the high-fidelity model in its domain, boundary conditions, and reaction coefficient; then a mixed-fidelity model adaptively formed based on $d_1$ might not give a low QoI error when used to infer from $d_2$. However, when the adaptivity algorithm is directly applied to this $d_2$, it is still able to find a mixed-fidelity model with a low QoI error. This behavior suggest that, if resources allow, it is best to perform the adaptivity online with the true observations. If one is confident that one's high-fidelity model is a good approximation of reality, then the online-offline approach can potentially give a faster yet accurate QoI estimate.

>>>>

The unknown parameters to be inferred correspond to the forcing field, so that in this case $f(q)=q$. The observations are sparse measurements of the state.
\cref{fig:baseSetup} shows the locations of the observations and the region $\Omega_I$ over which the QoI is calculated. For the low-fidelity model, the inverse problem is linear (i.e., the observations depend linearly on the parameters). Since the inverse problem is ill-posed, we use Tikhonov regularization~\cite{EngHanNeu00}; the regularization term in the objective function \cref{eq:invOpt_obj} is $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2\:\textrm{d}A$, where $\beta=10^{-5}$ is the regularization coefficient.
%
\begin{figure}[htbp]
\centering
\includegraphics[width=0.8\textwidth]{baseSeries/setup_3_3.pdf}
\caption{Locations of the observations and the QoI region.}
\label{fig:baseSetup}
\end{figure}
%

@@

We see that the largest local error contributions are concentrated in the QoI region and around the observation location closest to the QoI. In the first decomposition of the error (\cref{fig:baseRef0}), the region where the elemental error is greatest is around the leftmost observation location. Since the constraining model is an elliptic PDE, with weak convection, information flow is localized, and is weakly convected from left to right. Therefore, for the calculation of the QoI, it is most important to refine the region near the leftmost observation location, and then around the QoI region. After that, the error decomposition suggests refinement in regions upstream and around the middle observation location, and then the rightmost observation location.

@@

%------------------------------------------------------------%
\subsubsection{Interaction of Observations and QoI} \label{sec:qoivdata}
%------------------------------------------------------------%
%
The error estimate decomposition suggests the use of the high-fidelity model in areas of the domain that are important to the interaction between the observations and QoI; the interaction between these two can be complex, and the areas suggested for refinement may be nonintuitive. To see this, we compare the error estimate decomposition for three sizes of the QoI region $\Omega_I$ given the same set of observation locations, and for three nested sets of observation locations given the same QoI region. For the sake of illustration, we make two refinement iterations for each combination of observations and QoI region, regardless of the magnitude of the relative error estimate. However, in conducting the numerical experiments, it was observed that the number of iterations needed to achieve a given tolerance tended to increase as the QoI region increased.

\cref{fig:qoiStudy} shows the domain for three cases considered, with the same set of observations but increasingly large, nested QoI regions $\Omega_I$.
The error decompositions for each case are also shown in \cref{fig:qoiStudy}. The bottom row gives the baseline case presented in \cref{sec:cdvcdrBaseRef}, although here we choose the basis functions $i$ whose error $\varepsilon_i$ are among the largest $5\%$, rather than only enough basis functions to cover 5\% of the domain in their support, so the proportion of elements marked for refinement in each iteration will be slightly larger than in \cref{sec:cdvcdrBaseRef}.
% KW: I don't understand this last sentence. What's the difference with Section 4.1.2?
Although refinement is still most important around the observation location closest to $x_1=0$, as the QoI region expands the other two observation locations become more important in that the error decomposition suggests refinement around them earlier. As the QoI region expands, it is also more clearly noticeable that refinement is not equally important in all parts of the QoI region.

@@

\cref{fig:dataStudy} shows another set of cases considered, now with the same QoI region $\Omega_I$ but with increasing, nested sets of observations.
The error decomposition for the three cases is shown in \cref{fig:dataStudy}. The bottom row is the same as that in \cref{fig:qoiStudy}. Refinement appears to be consistently most important around the observation location closest to $x_1=0$ and the QoI region. However, as more observation locations are added, it becomes no longer necessarily true that refinement becomes less important around observation locations as their distance from the QoI region increases. In the second and third rows, one can see that after the areas around the QoI region and the two closest observation locations have been refined, the next area to be refined is not around the third closest observation location, but rather around one near the middle of the domain. This suggests that interactions between observation locations and the QoI may be non-intuitive, and in these cases a rigorous method for forming a mixed-fidelity model would be most helpful. % KW: I don't understand a lot of this discussion. You need to explain to me what you mean.

@@

We consider the same high-fidelity model as in \cref{sec:cdvcdr}:
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in U,
\end{equation}
with the same diffusion coefficient $k_d = 0.1$  and reaction coefficient $k_r = -42$. The low-fidelity model
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in\R
\end{equation}
differs from the high-fidelity model only in that the parameter $q$ is a constant instead of a field. The intermediate mixed-fidelity models thus have parameter fields that are constant over the subregions of the domain where the low-fidelity model is used. For ease of implementation, we require that the resulting parameter field remain continuous at the interface between the low-fidelity and high-fidelity subdomains, although this constraint is not necessary for the theory to hold. The domain, mesh, boundary conditions, and velocity field, as well as the observations, unknown parameters to be inferred, and QoI, remain the same as described in \cref{sec:cdvcdr}. As the inverse problem is ill-posed, except when the low-fidelity model is used throughout the domain, regularization is used; the Tikhonov regularization term in \cref{eq:invOpt_obj} is $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2+f(q)^2\:\textrm{d}A$, where $\beta=10^{-3}$ is a regularization coefficient.

@@

\Cref{fig:svfRef} shows the local error contributions, as well as the subdomains where the low- and high-fidelity models are used, for the first two and last mixed-fidelity model thus generated. 
%
\begin{figure}[htbp]
\centering
\subfloat[MF$_0$ ($0\%$ HF)]{
	\includegraphics[width=0.46\textwidth]{svf/cd_cdr_LF_divvy.png}
  \includegraphics[width=0.49\textwidth]{svf/err_breakdown_LF.png}
} \\
\subfloat[MF$_1$ ($10\%$ HF)]{
	\includegraphics[width=0.46\textwidth]{svf/cd_cdr_MF01_divvy.png}
  \includegraphics[width=0.49\textwidth]{svf/err_breakdown_MF01.png}
} \\
\subfloat[MF$_2$ ($20\%$ HF)]{
  \includegraphics[width=0.46\textwidth]{svf/cd_cdr_MF02_divvy.png}
  \includegraphics[width=0.49\textwidth]{svf/err_breakdown_MF02.png}
} \\
%\begin{subfigure}[b]{\textwidth}
%	\centering
%	\includegraphics[width=0.48\textwidth]{svf/cd_cdr_MF03_divvy.png}
%  \includegraphics[width=0.51\textwidth]{svf/err_breakdown_MF03.png}
%  \vspace{-0.7\baselineskip}
%  \caption{MF$_3$ ($30\%$ HF)}
%  \vspace{0.8\baselineskip}
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}
%	\centering
%	\includegraphics[width=0.48\textwidth]{svf/cd_cdr_MF04_divvy.png}
%  \includegraphics[width=0.51\textwidth]{svf/err_breakdown_MF04.png}
%  \vspace{-0.7\baselineskip}
%  \caption{MF$_4$ ($40\%$ HF)}
%  \vspace{0.8\baselineskip}
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}
%	\centering
%	\includegraphics[width=0.48\textwidth]{svf/cd_cdr_MF05_divvy.png}
%  \includegraphics[width=0.51\textwidth]{svf/err_breakdown_MF05.png}
%  \vspace{-0.7\baselineskip}
%  \caption{MF$_5$ ($50\%$ HF)}
%  \vspace{0.8\baselineskip}
%\end{subfigure}
\subfloat[MF$_6$ ($60\%$ HF)]{
	\includegraphics[width=0.46\textwidth]{svf/cd_cdr_MF06_divvy.png}
  \includegraphics[width=0.49\textwidth]{svf/err_breakdown_MF06.png}
}
\caption{Left: Multi-fidelity refinement over the domain (low-fidelity constant-parameter model used in red portion, high-fidelity field-parameter model used in blue portion). Right: local error contributions. The (weighted) residual, and thus the local error contribution, tends to spike sharply at the interface between the low- and high-fidelity regions; the color range is truncated to make the error distribution visible elsewhere in the domain.}
\label{fig:svfRef}
\end{figure}
%
Comparing to \cref{fig:baseRef}, we see that in this case the local error contribution is not as greatly concentrated around the QoI region and the nearest observation location; here, all three observation locations and the QoI region have associated regions of sufficiently similar high local error that all are refined in the first iteration. This reflects the global nature of the differences between the low- and high-fidelity models.

@@

The unknown parameters we wish to infer correspond to the source term $f(q)=q$; we impose $f(q)=q=0$ on the boundary $\partial\Omega$. Synthetic observations at 18 points in the domain are generated by running the high-fidelity model on a finer mesh. The locations of the observations as well as the QoI region $\Omega_I$ are shown in \cref{fig:setup3D}. We set the regularization term in \cref{eq:invOpt_obj} to be $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2\:\textrm{d}V$.
%
\begin{figure}[htbp]
\centering
\includegraphics[width=0.4\textwidth]{series3D/setup_aerial_nolegend.pdf} \hfill
\includegraphics[width=0.55\textwidth]{series3D/setup_3view.pdf} \\
\vspace{\baselineskip}
\includegraphics[width=0.6\textwidth]{series3D/setup_side_view.pdf}
\caption{Three views of the locations of the observations and the QoI region.}
\label{fig:setup3D}
\end{figure}
%

@@

We now present the results of solving the inference problem using \cref{alg:refSeries}, with a relative error tolerance of $0.1\%$. At each iteration, we choose the $2\%$ of the basis functions with the largest error for model refinement; since each linear Lagrange basis function has eight elements in its support, the number of additional elements marked for refinement in each iteration is usually greater than for the 2D examples. All simulations are run on a single processor; we use the default nonlinear solver in \texttt{libMesh} \cite{libMeshPaper} (Newton's method with Brent line-search), and linear solves are performed using PETSC's GMRES solver, preconditioned by incomplete factorization. % KW: need to add a citation to papers for libMesh and PETSC?

@@
